### 2016-11-23 R/ITHIM.Rmd
Now methods paper is created with knitr
See files ./lib/*.Rnw for editing content.

@@

DALYsTotal <- data.frame()
ITHIMList <- list()
ITHIM.default <- createITHIM()
ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
ITHIM.default <- updateITHIM(ITHIM.default, "cv", 1.65)
ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravel", 4)
ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)

@@

```{r ITHIMObjsQuintiles, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
DALYsTotal2 <- data.frame()
ITHIMList <- list()

@@

p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD") & type == "daly"), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( . ~ disease, scales = "free_y")
```

```{r newMETs, echo = TRUE, eval = FALSE, echo = FALSE}
newMETsFile <- "ATUS_to_METS_key.csv"

```{r readnewMETs, echo = TRUE, eval = FALSE, echo = FALSE}
cpsFile <- "~/GHI/data/ATUS/atuscps_2015.dat"
cps <- read.table(cpsFile, header=T, skip=0, sep="," )
cps <- within(cps, {
    TUCASEID <- as.character(TUCASEID)
    TULINENO <- as.character(TULINENO)
    GEREG <- as.factor(GEREG)
    })
cps <- subset(cps, TULINENO == "1")

```{r metKey, echo = TRUE, eval = FALSE, echo = FALSE}
allActivities <- colnames(summaryList$Data)
nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]

```{r parameters, eval = FALSE, echo = FALSE}
parameters <- ddply(D, .(SEX, AGECAT), summarise,
             mean = mean(METs, na.rm=T),
             sd = sd(METs, na.rm=T)
             )

(METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2)))

METmatrix <- METmatrix/METmatrix[3,2]

```{r doseresponse, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
RR1vec <- unique(unlist(createActiveTransportRRs()))
lapply(createActiveTransportRRs(),function(x) lapply(x, function(x) unique(c(x))))
#~~~~~~~~~~~~~~~~~~~~~~~
# Manual bullshit!
#~~~~~~~~~~~~~~~~~~~~~~~
RR1vec <- RR1vec[RR1vec != 1]
names(RR1vec) <- c("BreastCancer","ColonCancer.M","ColonCancer.F", "CVD", "Dementia", "Depression.young", "Depression.old", "Diabetes")

@@

>>>>

```{r newMETs, echo = TRUE, eval = TRUE, echo = FALSE}
newMETsFile <- "ATUS_to_METS_key.csv"

```{r readnewMETs, echo = TRUE, eval = TRUE, echo = FALSE}
cpsFile <- "~/GHI/data/ATUS/atuscps_2015.dat"
cps <- read.table(cpsFile, header=T, skip=0, sep="," )
cps <- within(cps, {
    TUCASEID <- as.character(TUCASEID)
    TULINENO <- as.character(TULINENO)
    GEREG <- as.factor(GEREG)
    })
cps <- subset(cps, TULINENO == "1")

```{r metKey, echo = TRUE, eval = TRUE, echo = FALSE}
allActivities <- colnames(summaryList$Data)
nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]

```{r parameters, eval = TRUE, echo = FALSE}
parameters <- ddply(D, .(SEX, AGECAT), summarise,
             mean = mean(METs, na.rm=T),
             sd = sd(METs, na.rm=T)
             )

METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))

METmatrix <- METmatrix/METmatrix[3,2]
METmatrix[2,] <- METmatrix[3,]
rownames(METmatrix) <- paste0("ageClass",1:8)
colnames(METmatrix) <- c("M","F")
METmatrix
```

@@

DALYsTotal <- data.frame()
ITHIMList <- list()
ITHIM.default <- createITHIM()
ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
ITHIM.default <- updateITHIM(ITHIM.default, "cv", 1.65)
ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravel", 4)
ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravelMatrix", METmatrix)
ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)

@@

@@

p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD") & type == "daly"), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( . ~ disease, scales = "free_y")
```

```{r doseresponse, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
RR1vec <- unique(unlist(createActiveTransportRRs()))
lapply(createActiveTransportRRs(),function(x) lapply(x, function(x) unique(c(x))))
#~~~~~~~~~~~~~~~~~~~~~~~
# Manual bull$%#@!
#~~~~~~~~~~~~~~~~~~~~~~~
RR1vec <- RR1vec[RR1vec != 1]
names(RR1vec) <- c("BreastCancer","ColonCancer.M","ColonCancer.F", "CVD", "Dementia", "Depression.young", "Depression.old", "Diabetes")

@@

### Appendix

```{r ITHIMObjsQuintiles, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
DALYsTotal2 <- data.frame()
ITHIMList <- list()

### 2016-11-23 methodsPaper/lib/materialsAndMethods.Rnw
Now methods paper is created with knitr
See files ./lib/*.Rnw for editing content.

As with the original ITHIM implementation we employed a comparative
risk analysis (CRA) across scenarios defined by mean active
transportation time, i.e., walking and cycling.  To do so we
investigate the proportional change in national disease burden between
the baseline model and a given alternative scenario, i.e., the
population attributable fraction, $\af$.  The population attributable
fraction is defined below and computed for each disease included in
the model, namely Breast Cancer, Colon Cancer, Depression, Dementia,
Diabetes and Cardiovascular disease.  For readability we omit an index
indicating disease.

\begin{equation}
\af = \frac{\int \! R(x)P(x) \, \mathrm{d}x  - \int \! R(x)Q(x) \,
  \mathrm{d}x}{\int \! R(x)P(x) \, \mathrm{d}x} \label{paf}
\end{equation}

Here $R(x)$ represents the risk when the individual has exposure $x$,
physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
density of individuals with exposure $x$, in the baseline and
alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of two independent random variables, travel-related
and non-travel-related physical activity.

@@

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Tue Nov  8 09:38:46 2016
\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & M & F \\
  \hline
0-4 & 0.00 & 0.00 \\
  5-14 & 0.00 & 0.00 \\
  15-29 & 0.97 & 1.00 \\
  30-44 & 1.04 & 1.12 \\
  45-59 & 0.95 & 0.99 \\
  60-69 & 0.90 & 0.94 \\
  70-79 & 0.83 & 0.88 \\
  80+ & 0.72 & 0.77 \\
   \hline
\end{tabular}
\end{table}

@@

@@

To estimate the disease burden we first approximate the population
attributable fraction through a discretization of the integral in
equation \ref{paf}.

\begin{equation}
\delta = 1 - \af \approx \frac{\sum_{i=1}^n R(x_i)}{\sum_{i=1}^n R(x_i^\prime)}
\end{equation}

The disease burden, in this case DALYs, is found for each of the
percentiles using equation \ref{burden}, using $B_0$, the overall nation-wide burden for
the baseline.  The overall burden is then computed as the sum across
quintiles, $B=\sum_{j=1}^n B_j$.

\begin{equation}
B_j = \delta B_0\frac{\RR_j}{\sum_{k=1}^n \RR_k}\label{burden}
\end{equation}

@@

\begin{equation}
\RR(x) = \mathrm{e}^{-r \tilde{x}}
\end{equation}

@@

\begin{figure}[t]
  \centerline{\includegraphics[width=\textwidth]{./figures/fig3.pdf}}
  \caption{Estimates for disease burden nationwide in terms of thousands of
    DALYs for each of the six diseases included in the model.
  }\label{fig3}
\end{figure}

>>>>

As with the original ITHIM implementation we employ a comparative risk
analysis (CRA) across scenarios defined by mean active transportation
time, i.e., walking and cycling.  To do so we estimate the
proportional change in disease burden between a baseline model and an
alternative scenario, namely the population attributable fraction,
$\af$.  The population attributable fraction is defined below and
computed for each disease included in the model.

\begin{equation}
\af_j = \frac{\int \! R_j(x)P(x) \, \mathrm{d}x  - \int \! R_j(x)Q(x) \,
  \mathrm{d}x}{\int \! R_j(x)P(x) \, \mathrm{d}x} \approx 1 - \frac{{\displaystyle \sum_{i=1}^n} R_j(x_i^\prime)}{{\displaystyle \sum_{i=1}^n} R_j(x_i)}  = 1 - \delta_j % = 1 - \frac{{\displaystyle \sum_{i=1}^n} e^{-\alpha \sqrt{x_i^\prime}}}{{\displaystyle \sum_{i=1}^n} e^{-\alpha \sqrt{x_i}}}
  \label{paf}
\end{equation}

Here $R(x)$ represents the risk when the individual has exposure $x$,
physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
density of individuals with exposure $x$, in the baseline and
alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of two independent random variables, travel-related
and non-travel-related physical activity.  $\mathbf{x}$ and
$\mathbf{x}^\prime$ represent the quantiles for the exposure
distribution in baseline and alternative scenarios, respectively.
$\delta_j$ is the proportional burden for disease $j$ in the alternate
scenario.

@@

@@

%To estimate the disease burden we first approximate the population
%attributable fraction through a discretization of the integral in
%equation \ref{paf}.

%% \begin{equation}
%% \delta = 1 - \af \approx \frac{{\displaystyle \sum_{i=1}^{n}} R(x_i)}{{\displaystyle \sum_{i=1}^{n}} R(x_i^\prime)}
%% \end{equation}

@@

The disease burden, in this case DALYs, is found for each of the
quantiles using equation \ref{burden} using $B_0$, the overall
nation-wide burden for the baseline.  The overall burden is then
computed as the sum across quintiles, $B=\sum_{j=1}^n B_j$.

\begin{equation}
B_j = \delta B_0\frac{\RR_j}{\sum_{k=1}^n \RR_j}\label{burden}
\end{equation}

@@

\begin{equation}
\RR_j(x) = \mathrm{e}^{-\alpha_j \tilde{x}}
\end{equation}

@@

\begin{sidewaysfigure}[t]
  \centerline{\includegraphics[width=\textwidth]{./figures/fig3.pdf}}
  \caption{Estimates for disease burden nationwide in terms of thousands of
    DALYs for each of the six diseases included in the model.
  }\label{fig3}
\end{sidewaysfigure}

### 2016-11-23 methodsPaper/tex/materialsAndMethods.Rnw
Now methods paper is created with knitr
See files ./lib/*.Rnw for editing content.

As with the original ITHIM implementation we employ a comparative risk
analysis (CRA) across scenarios defined by mean active transportation
time, i.e., walking and cycling.  To do so we estimate the
proportional change in disease burden between a baseline model and an
alternative scenario, namely the population attributable fraction,
$\af$.  The population attributable fraction is defined below and
computed for each disease included in the model.

>>>>

\Sexpr{1=1}
As with the original ITHIM implementation we employ a comparative risk
analysis (CRA) across scenarios defined by mean active transportation
time, i.e., walking and cycling.  To do so we estimate the
proportional change in disease burden between a baseline model and an
alternative scenario, namely the population attributable fraction,
$\af$.  The population attributable fraction is defined below and
computed for each disease included in the model.
